---
title: "How Linear Models Can Kill You"
author: "Tom Bruning"
date: "July 24, 2015"
output: ioslides_presentation
---

## The Problem

Whenever you look at linear models it is important to try to understand the underlying event behavior, not just the visible manifestation from time _0_ to _now_. 

The problem I have modeled is the trajectory of a missle from launch to the number of seconds that have elapsed since launch.  The only information available to the linear model is what has occurred, and that is incomplete information.

The following slides numerically show the behavior of the model.  The graphical model is available  [here](https://tgbruning.shinyapps.io/web_app).

## The Model

The formula used to compute the trajetory is:   
Height(_i_) =  100 ft + (200 ft * _i_)  - (5 * _i_^2), where _i_ = seconds since launch.

```{r,the_code, echo = FALSE, message= FALSE, warning=FALSE, fig.height= 3, fig.width= 5, results='asis'}
require(knitr)
height <- matrix(nrow = 40, ncol = 2)
h <- vector()
for( i in 1:40 ) {
    h[i] <- 100 + (200 * i)  - (5 * i^2);
    height[i,2] <- h[i]
    height[i,1] <- i
}
height_df <- as.data.frame(height)
colnames(height_df) <- c("i","height")
```

```{r}
height_df[1:5,]
```
---

### The Modeling Problem

As long as the missle ascends the linear model is adequate, but when gravity (represented by (5 * _i_^2) in the model overcomes the 200 ft. per second accention, the missle begins to fall.

```{r}
height_df[19:21,]
```
At this point the linear model is worthless, and in the case of the soldier standing at point 40,0 on the plot: fatal.

This model was taken from _How Not To Be Wrong: The Power of Mathamatical Thinking_ by Jordan Ellenberg.

--- 
### The Code

```{r, show_code, ref.label= "the_code" }
```